---
title: "Case_Study_4_Code"
author: "Brian Gaither, Sean Mcwhirter, Andrew Mejia, Sabrina Purvis"
date: "`r Sys.time()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 6
  github_document:
    toc: yes
    toc_depth: 6
  word_document:
    toc: yes
    toc_depth: '6'
always_allow_html: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(hms)
library(rvest)
library(lubridate)
library(foreach)
library(stringr)
library(iterators)
library(progress)
library(doParallel)
library(doSNOW)
library(dplyr)
library(states)
library(ggplot2)
library(ggthemes)
```

```{r}
years = c(1999:2012)
division = 'Overall+Women'
section = '10M' 
sex = 'W'
```

#### Functions

#The gen_Link function will generates the link with the query parameters for the searchable database

```{r}
gen_Link = function(year,division,section,page=1,sex){

  paste0( 'http://www.cballtimeresults.org/performances'
          ,'?division=',division,'&page=',page,
          '&section=',section, '&sex=',sex, 
          '&utf8=%E2%9C%93','&year=',year)
          #,'?utf8=%E2%9C%93&section=',section
          #,'&year=',year,'&division=',division,'&page=', page)

}

```

# The gen_Table function will parse through the table of 20 records and 15 observations
# and parse the table into its own data frame from xml2 read)html function and then use 
# the pipe operator to use rvest nodes to find the table structure 
# then the fucntion will insert the metadata for the query parameters of year, division, section, page, source link and sex 
```{r}
gen_Table = function(year,division,section,page, sex){
  
  
  #use gen_link function to get link to page 
  genlink=gen_Link(year,division,section,page=page, sex=sex)
  
  #read the page, and grab to 'table' tag
  single_table = xml2::read_html(genlink) %>% 
    rvest::html_nodes("table")  %>% 
    rvest::html_table(fill=TRUE) 
  
  #get the table and add metadata for the query parameters
  table_out = single_table[[1]] %>% 
    mutate(year=year, divisionTitle=division, section=section, page=page, source=genlink, sex = sex)
  
}

```

# This function will use all available cores on machine
# It will process the years in parrell. 
# This code has been adapted from 
# https://github.com/ngupta23/ds7333_qtw/blob/master/case_study_2/submission_Kannan_Moro_Gupta/code/CS2_ETL.Rmd 
# https://cran.r-project.org/web/packages/doSNOW/doSNOW.pdf
# https://stackoverflow.com/questions/36794063/r-foreach-from-single-machine-to-cluster
# https://cran.r-project.org/web/packages/progress/progress.pdf
# https://www.r-bloggers.com/2013/08/the-wonders-of-foreach/


```{r}
scrapeTables  = function(years,division,section, sex,  max_itr = 500){
  
  
library(progress)
library(doParallel)
library(doSNOW)  
  
    #Initialize Parrellel Process to detect number of cores 
    #https://cran.r-project.org/web/packages/doSNOW/doSNOW.pdf
    #https://stackoverflow.com/questions/36794063/r-foreach-from-single-machine-to-cluster
  
    #Generate and register initial clusters based on cores, otherwise, this is a long process 
    cl = makeCluster(detectCores())
    doSNOW::registerDoSNOW(cl)
    
    #Generate progress bar for the parallel loop based on number of years 
    #https://cran.r-project.org/web/packages/progress/progress.pdf
    progBar = progress::progress_bar$new(total = length(years),format='[:bar] :percent :eta')
    progress = function(n) progBar$tick()
    
    #Initialize a parallel loop per each year
    #Intialize tableRaw as empty to loop to be populated as a table dataframe from gen_Table function
    tableRaw=NULL
    
    #tableRaw will now use for each using years as the iterator to use .combine to rbind
    #.export will do the gen_Table and gen_Link functions simultanlously and 
    #options.snow will show the progress bar
    # the %dopar% will process all the years simultaneously. 
    #https://www.r-bloggers.com/2013/08/the-wonders-of-foreach/
    tableRaw = foreach(y=years
                      ,.combine=rbind,.export=c('gen_Table','gen_Link')
                      ,.options.snow = list(progress=progress)) %dopar%
      {
        
        library(foreach)
        library(dplyr)
        #intialize isCompleted variable as FALSE for bool conditions to see if loop has been completed
        isCompleted=FALSE
        
        #Initiate loop since most pages are 487, we will only loop for the iterations for max_itr
        tableRaw=foreach(p=c(1:max_itr),.combine=rbind) %do% 
          if(!isCompleted) {
            message('getting year:',y, ' page:',p,appendLF = F)
            #get the table of the current page
            table = gen_Table(year=y
                             ,division=division
                             ,section=section
                             ,page=p
                             ,sex=sex)
            message(' rows:',nrow(table))
            isCompleted = nrow(table)==0 #if there is record, we are at the last page, no need to read further
            return(table)
          }
        return(tableRaw)
      }
    #Deactivate the cluster of cores
    stopCluster(cl)
    #save the raw data to rda format for later processing based on gender
    saveRDS(tableRaw,file=paste0('CB',sex,'tableRaw.rds'))
    
  return(tableRaw)
}
  
 
```

### The purpose of this function is to transform the raw tables from the scrape ###
### https://stackoverflow.com/questions/50040968/convert-a-duration-hms-to-seconds ### 
### https://stackoverflow.com/questions/10835908/is-there-a-way-to-convert-mmss-00-to-seconds-00 ### 
### https://stackoverflow.com/questions/24173194/remove-parentheses-and-text-within-from-strings-in-r  ###

```{r}
tableTransform =function(data_df, cols_to_remove=NULL){
  dataDF = data_df %>% 
    #Seperate Home town into seperate columns
    separate(col = 'Hometown', c('HomeTown', 'HomeState'), sep = ',', extra = 'merge', remove = TRUE, fill = 'right') %>% 
    #Seperate PiS/TiS into seperate columns 
    separate(col='PiS/TiS',c('PiS','TiS'),sep='\\/'
             ,extra='drop',remove=TRUE) %>% 
    #Seperate PiD/TiD into seperate columns 
    separate(col='PiD/TiD',c('PiD','TiD'),sep='\\/'
             ,extra='drop',remove=TRUE) %>% 
    #Trim the casted upper HomeTown strings of whitespace 
    mutate(HomeTown = toupper(trimws(HomeTown))
           , HomeState = toupper(trimws(HomeState))
           , HomeTown = ifelse(HomeTown %in% c('NR', '', NULL), NA, toupper(trimws(HomeTown)))
           , HomeState = ifelse(HomeState %in% c('NR', '', NULL), NA, toupper(trimws(HomeState)))
           #Check if HomeState is in state.abb or DC and return USA else return the HomeTown as the Country
           , HomeCountry = ifelse(HomeState %in% c(state.abb, "DC"), "USA", HomeTown)
           #Remove White Space
           , PiS = ifelse(trimws(PiS) %in% c('NR', '', NULL), NA, trimws(PiS))
           , TiS = ifelse(trimws(TiS) %in% c('NR', '', NULL), NA, trimws(TiS))
           , PiD = ifelse(trimws(PiD) %in% c('NR', '', NULL), NA, trimws(PiD))
           , TiD = ifelse(trimws(TiD) %in% c('NR', '', NULL), NA, trimws(TiD))
           , Division = ifelse(trimws(Division) %in% c('NR', '', NULL), NA, trimws(Division))
           # Normalize Time to seconds and minutes 
           , RawTime = strptime(Time, format='%H:%M:%S')
           , RawTime_S = RawTime$hour * 3600 + RawTime$min * 60 + RawTime$sec 
           , RawTime_M = as.numeric(RawTime_S)/60
           , RawPace = strptime(Pace, format = "%M:%OS")
           , RawPace_S = RawPace$min * 60 + RawPace$sec 
           #Normalize Age where 'NR' as NA 
           , Age = ifelse(Age %in% c("NR"), NA, Age)
           #Cast Variables as appropriate dtypes 
           , Age = as.numeric(Age)
           , RawTime_S = as.numeric(RawTime_S)
           , RawPace_S = as.numeric(RawPace_S)
           , RawPace_M = as.numeric(RawPace_S)/60
           , year = as.factor(year)
           #Remove the (<Sex>) from the names
           #, Name = str_replace(Name, " \\s*\\([^\\)]+\\)", '')
           )
  #Remove columns we do not want 
  dataDF = dataDF %>% select (-all_of(c(cols_to_remove)))
                  
  return(dataDF %>% select(c( Age, year,  HomeTown, HomeState, HomeCountry, RawTime_S, RawTime_M, RawPace_S, RawPace_M, sex, section, PiS, TiS, PiD, TiD)))
}
```



### Perform Scrape ### 

### Caution takes a long time without hexacore machine ### 
```{r}
#Women_table = scrapeTables(years=years,division = division,section=section,sex = sex, max_itr = 500)
```

```{r}
#men_table = scrapeTables(years=years,division = 'Overall+Men',section=section,sex = "M", max_itr = 500)

```

### Load tables from file ### 

```{r}
mens_table <- readRDS("/media/andrew/Seagate Backup Plus Drive/Documents/School/HomeWork/QTW/DS7333/CASE_STUDY_4/CBMtableRaw.rds")

womens_table <- readRDS("/media/andrew/Seagate Backup Plus Drive/Documents/School/HomeWork/QTW/DS7333/CASE_STUDY_4/CBWtableRaw.rds")
```

```{r}
view(head(womens_table, n = 10))
```

```{r}
cols_for_remove = c("divisionTitle", "source", "Pace", "Time", "RawTime", "RawPace", "page", "Race", "source")
mens_table_T = tableTransform(mens_table, cols_to_remove = all_of(cols_for_remove))
womens_table_T = tableTransform(womens_table, cols_to_remove = all_of(cols_for_remove))

```

```{r}
head(womens_table_T, n = 10)
```

```{r}
summary(womens_table_T)
```

```{r}
columns = c("Age", "year",  "HomeTown", "HomeState", "HomeCountry", "RawTime_S", "RawTime_M", "RawPace_S", "RawPace_M", "sex", "section", "PiS", "TiS", "PiD", "TiD")

womens_table_T_na = womens_table_T %>% filter_at(vars(all_of(columns)),any_vars(is.na(.)))

womens_table_T_na
```

```{r}
na_df = data.frame(rowsum(+(is.na(womens_table_T)), womens_table_T$year))
na_df = cbind(year_idx = rownames(na_df), na_df)
```


```{r}
p1_na = ggplot(na_df, aes(x=year_idx, y=HomeTown)) + geom_bar(stat = "identity", fill = "grey") + labs(title = "Counts of missing HomeTown Records by Year") + xlab("Year Index") + ylab("Count of HomeTown")
p1_na
```

```{r}
p2_na = ggplot(na_df, aes(x=year_idx, y=HomeState)) + geom_bar(stat = "identity", fill = "grey") +  labs(title = "Counts of missing HomeState Records by Year") + xlab("Year Index") + ylab("Count of HomeState")

p2_na
```

## We will remove NAs and only focus on completed records and rewmove NAs for further processing and reindex the table. 
```{r}
womens_table_T = womens_table_T[complete.cases(womens_table_T), ]
row.names(womens_table_T) = NULL 
```

```{r}
head(womens_table_T, n = 10)
```
```{r}
womensTableInfo = data.frame(Variable = names(womens_table_T),Type = sapply(womens_table_T, typeof), row.names = NULL)
womensTableInfo
```

#############################################################################################################################################################################################################################################################################################################################################################################################################################################################################################

```{r}
plot_data_age = ggplot(womens_table_T, aes(x = year, y = Age)) + geom_boxplot() + labs(title = "Distribution of Women Participants Age by Year") + xlab("Year") + ylab("Age")

plot_data_age
```

```{r}
plot_data_time = ggplot(womens_table_T, aes(x = year, y = RawTime_M)) + geom_boxplot() + labs(title = "Distribution of Women Raw Time by Year") + xlab("Year") + ylab("Minutes")

plot_data_time
```

```{r}
plot_data_Pace = ggplot(womens_table_T, aes(x = year, y = RawPace_M)) + geom_boxplot() + labs(title = "Distribution of Women Raw Pace by Year") + xlab("Year") + ylab("Minutes")


plot_data_Pace
```