---
title: "Case_Study_4_Code"
author: "Brian Gaither, Sean Mcwhirter, Andrew Mejia, Sabrina Purvis"
date: "`r Sys.time()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 6
  github_document:
    toc: yes
    toc_depth: 6
  word_document:
    toc: yes
    toc_depth: '6'
always_allow_html: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(hms)
library(rvest)
library(lubridate)
library(foreach)
library(stringr)
library(iterators)
library(progress)
library(doParallel)
library(doSNOW)
library(dplyr)
library(states)
library(ggplot2)
library(ggthemes)
```

```{r}
years = c(1999:2012)
division = 'Overall+Women'
section = '10M' 
sex = 'W'
```

#### Functions

#The gen_Link function will generates the link with the query parameters for the searchable database

```{r}
gen_Link = function(year,division,section,page=1,sex){

  paste0( 'http://www.cballtimeresults.org/performances'
          ,'?division=',division,'&page=',page,
          '&section=',section, '&sex=',sex, 
          '&utf8=%E2%9C%93','&year=',year)
          #,'?utf8=%E2%9C%93&section=',section
          #,'&year=',year,'&division=',division,'&page=', page)

}

```

# The gen_Table function will parse through the table of 20 records and 15 observations
# and parse the table into its own data frame from xml2 read)html function and then use 
# the pipe operator to use rvest nodes to find the table structure 
# then the fucntion will insert the metadata for the query parameters of year, division, section, page, source link and sex 
```{r}
gen_Table = function(year,division,section,page, sex){
  
  
  #use gen_link function to get link to page 
  genlink=gen_Link(year,division,section,page=page, sex=sex)
  
  #read the page, and grab to 'table' tag
  single_table = xml2::read_html(genlink) %>% 
    rvest::html_nodes("table")  %>% 
    rvest::html_table(fill=TRUE) 
  
  #get the table and add metadata for the query parameters
  table_out = single_table[[1]] %>% 
    mutate(year=year, divisionTitle=division, section=section, page=page, source=genlink, sex = sex)
  
}

```

# This function will use all available cores on machine
# It will process the years in parrell. 
# This code has been adapted from 
# https://github.com/ngupta23/ds7333_qtw/blob/master/case_study_2/submission_Kannan_Moro_Gupta/code/CS2_ETL.Rmd 
# https://cran.r-project.org/web/packages/doSNOW/doSNOW.pdf
# https://stackoverflow.com/questions/36794063/r-foreach-from-single-machine-to-cluster
# https://cran.r-project.org/web/packages/progress/progress.pdf
# https://www.r-bloggers.com/2013/08/the-wonders-of-foreach/


```{r}
scrapeTables  = function(years,division,section, sex,  max_itr = 500){
  
  
library(progress)
library(doParallel)
library(doSNOW)  
  
    #Initialize Parrellel Process to detect number of cores 
    #https://cran.r-project.org/web/packages/doSNOW/doSNOW.pdf
    #https://stackoverflow.com/questions/36794063/r-foreach-from-single-machine-to-cluster
  
    #Generate and register initial clusters based on cores, otherwise, this is a long process 
    cl = makeCluster(detectCores())
    doSNOW::registerDoSNOW(cl)
    
    #Generate progress bar for the parallel loop based on number of years 
    #https://cran.r-project.org/web/packages/progress/progress.pdf
    progBar = progress::progress_bar$new(total = length(years),format='[:bar] :percent :eta')
    progress = function(n) progBar$tick()
    
    #Initialize a parallel loop per each year
    #Intialize tableRaw as empty to loop to be populated as a table dataframe from gen_Table function
    tableRaw=NULL
    
    #tableRaw will now use for each using years as the iterator to use .combine to rbind
    #.export will do the gen_Table and gen_Link functions simultanlously and 
    #options.snow will show the progress bar
    # the %dopar% will process all the years simultaneously. 
    #https://www.r-bloggers.com/2013/08/the-wonders-of-foreach/
    tableRaw = foreach(y=years
                      ,.combine=rbind,.export=c('gen_Table','gen_Link')
                      ,.options.snow = list(progress=progress)) %dopar%
      {
        
        library(foreach)
        library(dplyr)
        #intialize isCompleted variable as FALSE for bool conditions to see if loop has been completed
        isCompleted=FALSE
        
        #Initiate loop since most pages are 487, we will only loop for the iterations for max_itr
        tableRaw=foreach(p=c(1:max_itr),.combine=rbind) %do% 
          if(!isCompleted) {
            message('getting year:',y, ' page:',p,appendLF = F)
            #get the table of the current page
            table = gen_Table(year=y
                             ,division=division
                             ,section=section
                             ,page=p
                             ,sex=sex)
            message(' rows:',nrow(table))
            isCompleted = nrow(table)==0 #if there is record, we are at the last page, no need to read further
            return(table)
          }
        return(tableRaw)
      }
    #Deactivate the cluster of cores
    stopCluster(cl)
    #save the raw data to rda format for later processing based on gender
    saveRDS(tableRaw,file=paste0('CB',sex,'tableRaw.rds'))
    
  return(tableRaw)
}
  
 
```

### The purpose of this function is to transform the raw tables from the scrape ###
### https://stackoverflow.com/questions/50040968/convert-a-duration-hms-to-seconds ### 
### https://stackoverflow.com/questions/10835908/is-there-a-way-to-convert-mmss-00-to-seconds-00 ### 
### https://stackoverflow.com/questions/24173194/remove-parentheses-and-text-within-from-strings-in-r  ###

```{r}
tableTransform =function(data_df, cols_to_remove=NULL){
  dataDF = data_df %>% 
    #Seperate Home town into seperate columns
    separate(col = 'Hometown', c('HomeTown', 'HomeState'), sep = ',', extra = 'merge', remove = TRUE, fill = 'right') %>% 
    #Seperate PiS/TiS into seperate columns 
    separate(col='PiS/TiS',c('PiS','TiS'),sep='\\/'
             ,extra='drop',remove=TRUE) %>% 
    #Seperate PiD/TiD into seperate columns 
    separate(col='PiD/TiD',c('PiD','TiD'),sep='\\/'
             ,extra='drop',remove=TRUE) %>% 
    #Trim the casted upper HomeTown strings of whitespace 
    mutate(HomeTown = toupper(trimws(HomeTown))
           , HomeState = toupper(trimws(HomeState))
           , HomeTown = ifelse(HomeTown %in% c('NR', '', NULL), NA, toupper(trimws(HomeTown)))
           , HomeState = ifelse(HomeState %in% c('NR', '', NULL), NA, toupper(trimws(HomeState)))
           #Check if HomeState is in state.abb or DC and return USA else return the HomeTown as the Country
           , HomeCountry = ifelse(HomeState %in% c(state.abb, "DC"), "USA", HomeTown)
           #Remove White Space
           , PiS = ifelse(trimws(PiS) %in% c('NR', '', NULL), NA, trimws(PiS))
           , TiS = ifelse(trimws(TiS) %in% c('NR', '', NULL), NA, trimws(TiS))
           , PiD = ifelse(trimws(PiD) %in% c('NR', '', NULL), NA, trimws(PiD))
           , TiD = ifelse(trimws(TiD) %in% c('NR', '', NULL), NA, trimws(TiD))
           , Division = ifelse(trimws(Division) %in% c('NR', '', NULL), NA, trimws(Division))
           # Normalize Time to seconds and minutes 
           , RawTime = strptime(Time, format='%H:%M:%S')
           , RawTime_S = RawTime$hour * 3600 + RawTime$min * 60 + RawTime$sec 
           , RawTime_M = as.numeric(RawTime_S)/60
           , RawPace = strptime(Pace, format = "%M:%OS")
           , RawPace_S = RawPace$min * 60 + RawPace$sec 
           #Normalize Age where 'NR' as NA 
           , Age = ifelse(Age %in% c("NR"), NA, Age)
           #Cast Variables as appropriate dtypes 
           , Age = as.numeric(Age)
           , RawTime_S = as.numeric(RawTime_S)
           , RawPace_S = as.numeric(RawPace_S)
           , RawPace_M = as.numeric(RawPace_S)/60
           , year = as.factor(year)
           #Remove the (<Sex>) from the names
           #, Name = str_replace(Name, " \\s*\\([^\\)]+\\)", '')
           )
  #Remove columns we do not want 
  dataDF = dataDF %>% select (-all_of(c(cols_to_remove)))
                  
  return(dataDF %>% select(c( Age, year,  HomeTown, HomeState, HomeCountry, RawTime_S, RawTime_M, RawPace_S, RawPace_M, sex, section, PiS, TiS, PiD, TiD)))
}
```



### Perform Scrape ### 

### Caution takes a long time without hexacore machine ### 
```{r}
#Women_table = scrapeTables(years=years,division = division,section=section,sex = sex, max_itr = 500)
```

```{r}
#men_table = scrapeTables(years=years,division = 'Overall+Men',section=section,sex = "M", max_itr = 500)

```

### Load tables from file ### 

```{r}
mens_table <- readRDS("/media/andrew/Seagate Backup Plus Drive/Documents/School/HomeWork/QTW/DS7333/CASE_STUDY_4/CBMtableRaw.rds")

womens_table <- readRDS("/media/andrew/Seagate Backup Plus Drive/Documents/School/HomeWork/QTW/DS7333/CASE_STUDY_4/CBWtableRaw.rds")
```

```{r}
head(womens_table, n = 10)
```

```{r}
cols_for_remove = c("divisionTitle", "source", "Pace", "Time", "RawTime", "RawPace", "page", "Race", "source")
mens_table_T = tableTransform(mens_table, cols_to_remove = all_of(cols_for_remove))
womens_table_T = tableTransform(womens_table, cols_to_remove = all_of(cols_for_remove))

```

```{r}
head(womens_table_T, n = 10)
dim(womens_table_T)
```

```{r}
summary(womens_table_T)
```

```{r}
columns = c("Age", "year",  "HomeTown", "HomeState", "HomeCountry", "RawTime_S", "RawTime_M", "RawPace_S", "RawPace_M", "sex", "section", "PiS", "TiS", "PiD", "TiD")

womens_table_T_na = womens_table_T %>% filter_at(vars(all_of(columns)),any_vars(is.na(.)))

head(womens_table_T_na, n = 10) 
dim(womens_table_T_na)
```

```{r}
na_df = data.frame(rowsum(+(is.na(womens_table_T)), womens_table_T$year))
na_df = cbind(year_idx = rownames(na_df), na_df)
```


```{r}
p1_na = ggplot(na_df, aes(x=year_idx, y=HomeTown)) + geom_bar(stat = "identity", fill = "grey") + labs(title = "Counts of missing HomeTown Records by Year") + xlab("Year Index") + ylab("Count of HomeTown")
p1_na
```

```{r}
p2_na = ggplot(na_df, aes(x=year_idx, y=HomeState)) + geom_bar(stat = "identity", fill = "grey") +  labs(title = "Counts of missing HomeState Records by Year") + xlab("Year Index") + ylab("Count of HomeState")

p2_na
```

## We will remove NAs and only focus on completed records and rewmove NAs for further processing and reindex the table. 
```{r}
womens_table_T = womens_table_T[complete.cases(womens_table_T), ]
row.names(womens_table_T) = NULL 
```

```{r}
head(womens_table_T, n = 10)
```
```{r}
Descriptions = c('Participant Age', 'Year of race', 'Participant Home Town', 'Participant Home State', 'Participant Home Country', "Participant's Time in Seconds", "Participant's Time in Minutes", "Participants Mile Pace in Seconds", "Participants Mile Pace in Minutes", "Gender", "Participant Race", "Position in Sex", "Total in Sex", "Position in Division", "Total in Division")
womensTableInfo = data.frame(Feature = names(womens_table_T), Description = Descriptions , Type = sapply(womens_table_T, typeof), row.names = NULL)
womensTableInfo
```

#############################################################################################################################################################################################################################################################################################################################################################################################################################################################################################

```{r}
plot_data_age = ggplot(womens_table_T, aes(x = year, y = Age)) + geom_boxplot() + labs(title = "Distribution of Women Participants Age by Year") + xlab("Year") + ylab("Age")

plot_data_age
```

```{r}
plot_data_time = ggplot(womens_table_T, aes(x = year, y = RawTime_M)) + geom_boxplot() + labs(title = "Distribution of Women Raw Time by Year") + xlab("Year") + ylab("Minutes")

plot_data_time
```

```{r}
plot_data_Pace = ggplot(womens_table_T, aes(x = year, y = RawPace_M)) + geom_boxplot() + labs(title = "Distribution of Women Raw Pace by Year") + xlab("Year") + ylab("Minutes")


plot_data_Pace
```

##########################################################################################################################################################################################################################################################################################################################################################
BRIAN GAITHER
#############################################################################################################################################################################
#############################################################################################################################################################################

Creating an age bin column for analysis
```{r}
womens_table_T$AgeBin = cut(womens_table_T$Age, breaks=c(0,5,15,25,35,45,55,65,75,85,95),labels=c("1-5","6-15","16-25","26-35","36-45","46-55","56-65","66-75","76-85","86-95"))
```

Creating a column to bin the pace in minutes for later analysis
```{r}
womens_table_T$PaceBin = cut(womens_table_T$RawPace_M, breaks=c(0,5,5.5,6,6.5,7,7.5,8,8.5,9,9.5,10,10.5,11,11.5,12,12.5,13,13.5,14,18), 
                 labels=c("1-5","5.1-5.5","5.5-6","6-6.5","6.5-7","7-7.5","7.5-8","8-8.5","8.5-9","9-9.5","9.5-10","10-10.5","10.5-11","11-11.5","11.5-12","12-12.5","12.5-13","13-13.5","13.5-14","11-18"))

```

write out the dataframe for later retrieval
```{r}
#write.csv(womens_table_T,"C:/Users/blgai/OneDrive/Documents/School/SMU/Spring 2021/Quantifying the World/Case4/files/women_data.csv", row.names = FALSE)
```

read the data into dataframe
```{r}
#womens_table_T = read.csv("C:/Users/blgai/OneDrive/Documents/School/SMU/Spring 2021/Quantifying the World/Case4/files/women_data.csv", header=TRUE)
```

Let's convert some of the data types to make the data easier to work with later
```{r}
womens_table_T = transform(womens_table_T, Race = as.factor(Race),Age = as.integer(Age), HTown = as.factor(HTown), HState = as.factor(HState), HCountry = as.factor(HCountry), year = as.factor(year) )
```


```{r}
ggplot(womens_table_T, aes(x = Age, y=RawTime_M)) + 
  geom_hex() +
  theme_bw() + 
  labs(title="Scatter Plot of Female Runners:  Age vs Time in Minutes")

```
```{r}
f = ggplot(womens_table_T)
f + geom_boxplot(mapping = aes(x=AgeBin, y=RawTime_M)) + theme_classic() + labs(title="Boxplot of Age Group by Running Time (Minutes)")
```

Let's examine overall runner volume
```{r}
dfRunVol = womens_table_T %>% group_by(year) %>% tally()
dfRunVol$year = as.integer(dfRunVol$year)
names(dfRunVol)[2] <- "RunnerVolume"
head(dfRunVol)
```

Plotting out volume of runners by year
```{r}
ggplot(dfRunVol, aes(x=year, y=RunnerVolume)) +
  geom_col() +
  labs(title="Volume of runners by year") +
geom_text(aes(label = RunnerVolume), vjust = -0.5)
```
let's see how well we can predict the volume of racers in the 15th year
```{r}


lmRunVol = lm(RunnerVolume ~ year, data = dfRunVol)
lmRunVol$coefficients

```
Below we see we have a statistically significant slope and intercept
```{r}
summary(lmRunVol)
```

See how well our line fits
```{r}
preds = predict(lmRunVol)
ggplot(dfRunVol, aes(x=year, y=RunnerVolume)) + geom_point() + geom_line(dfRunVol, mapping = aes(x=year, y=preds, col = "red")) + ggtitle("Prediction of Runner Volume")
```


How many runners are predicted to attend in 2013 (year 15)
```{r}
y_hat = 800.5 + (15*613.4)

print(paste0("Predicted volume of runners in 2013 is: ", y_hat))
```


95% confidence intervals
```{r}
dfNew = data.frame(year=15)
predict(lmRunVol, newdata = dfNew, interval = 'confidence')
```

we can see below that the IQR of running times throughout the years has remained relatively constant.  That means that
```{r}
f = ggplot(womens_table_T)
f + geom_boxplot(mapping = aes(x=year, y=RawTime_M)) + theme_classic() + labs(title="Boxplot of Running Time (Minutes) by Year")
```


Let's look at the volume of runners by pace bin.  We can see that the majority of runners are between 8-11 minute pace.  This means that a large swell of runners will be running through the course together and race support must have enough volunteers to support the increase of runners as they make their way through the course.
```{r}
womens_table_T$PaceBin <- as.character(womens_table_T$PaceBin)
womens_table_T$PaceBin <- factor(womens_table_T$PaceBin, levels=unique(womens_table_T$PaceBin))

ggplot(womens_table_T %>% group_by(year, PaceBin) %>% tally()) +
  geom_col(mapping = aes(x=PaceBin, y=n)) +
  facet_grid(vars(year)) +
  labs(title="Distribution of pace throughout the years")

```

Let's take a look at the median pace per year.
```{r}
womens_table_T %>% group_by(year) %>% dplyr::summarise(median_pace = median(RawPace_M)) %>% ggplot(aes(x=year, y=median_pace)) + geom_point() + ggtitle("Median Pace Per Year")
```

If we look at the max times throughout the years, there isn't a consistent pattern observable, we'll revisit this after looking more at the age bins further
```{r}
womens_table_T %>% group_by(year) %>% dplyr::summarise(max_time = max( RawTime_M )) %>% ggplot(aes(x=year, y=max_time)) + geom_point()
```

we see below that there has been a steady increase in 26-35 year old runners over the years
```{r}
ggplot(womens_table_T %>% group_by(year, AgeBin) %>% tally()) +
  geom_col(mapping = aes(x=AgeBin, y=n)) +
  facet_grid(vars(year)) +
  labs(title="Distribution of age throughout the years")

```
Let's take a closer look at the volume of runners in the 26-35 age category
```{r}
dfRunAgeBinVol = womens_table_T %>% group_by(year, AgeBin) %>% dplyr::summarise(AgeBinVolume = n()) 
head(dfRunAgeBinVol)
ggplot(dfRunAgeBinVol[dfRunAgeBinVol$AgeBin == '26-35',]) +
  geom_point(mapping = aes(x=year, y=AgeBinVolume)) +
  #facet_wrap(vars(year)) +
  labs(title="Increase in volume of 26-35 age group throughout the years")
```

```{r}
dfRunAgeBinVol$year = as.numeric(as.character(dfRunAgeBinVol$year))
dfRunAgeBinVol[dfRunAgeBinVol$AgeBin == '26-35',]
```
Let's trend the volume of 26-35 year old runners using LOESS
```{r}
df26 = dfRunAgeBinVol[dfRunAgeBinVol$AgeBin == '26-35',]
loessAge26 = loess(AgeBinVolume ~year, data=df26)
preds = predict(loessAge26)
ggplot(df26, aes(x=year, y=AgeBinVolume)) + 
  geom_point() + 
  geom_line(df26, mapping = aes(x=year, y=preds, col = "red")) + 
  ggtitle("Prediction of Runner Volume in 26-35 Age Category")
```
Let's now try to predict the volume of 26-35 year olds using a piece wise linear regression model
```{r}

pw.model = piecewise.linear(df26$year, df26$AgeBinVolume, middle=1, CI=TRUE, sig.level = 0.05)
pw.model
```

Now that we have a piece wise linear model, let's predict the volume of 26-35 year olds in years 2013 and 2014
```{r}
preds = predict(pw.model,c(1999:2014))
predYears = c(1999:2014)
dfPreds = data.frame(year = predYears, predAgeBinVolume = predict(pw.model,predYears))
ggplot(df26, aes(x=year, y=AgeBinVolume)) + 
  geom_point() + 
  geom_line(dfPreds, mapping = aes(x=year, y=predAgeBinVolume, col = "red")) + 
  geom_point(dfPreds[dfPreds$year %in% c(2013:2014), ], mapping = aes(x=year, y=predAgeBinVolume, col = "blue")) +
  ggtitle("Prediction of Runner Volume in 26-35 Age Category in years 2013 and 2014")
```

The actual predicted values for years 2013 and 2014.  This is what we want to share with the race management team
```{r}
predict(pw.model,c(2013:2014))
```

Let's look at average times of age 25-36 year olds
```{r}
dfAvgTimes26 = womens_table_T %>% group_by(year, AgeBin) %>% dplyr::summarise(avgTime = mean(RawTime_M))
dfAvgTimes26 = dfAvgTimes26[dfAvgTimes26$AgeBin == '26-35',]
dfAvgTimes26 %>% ggplot(aes(x=year, y=avgTime/60)) + geom_point() + ggtitle("Average Race Time Per Year for 26-35 Age Category") + labs(y="Race Time (hr)")
```

let's get a look at the pace of 26-35 year olds.  over the years, they've held a pretty consistent pace.  50% are in the 9-10 minute pace group.
```{r}
f = ggplot(womens_table_T %>% filter(womens_table_T$AgeBin == "26-35"))
f + geom_boxplot(mapping = aes(x=year, y=RawPace_M)) + theme_classic() + labs(title="Boxplot of Pace (Minutes) by Year for 26-35 aged runners")
```

Let's look at median times of age 25-36 year olds
```{r}
dfMedTimes26 = womens_table_T %>% group_by(year, AgeBin) %>% dplyr::summarise(medTime = median(RawTime_M))
dfMedTimes26 = dfMedTimes26[dfMedTimes26$AgeBin == '26-35',]
dfMedTimes26 %>% ggplot(aes(x=year, y=medTime/60)) + geom_point() + ggtitle("Median Race Time Per Year for 26-35 Age Category") + labs(y="Race Time (hr)")  +
geom_text(aes(label = medTime), vjust = -0.5)
```



Let's look at the proportion the 26-35 year olds make up of the total runners over the years. As of 2012, this age group makes up about 50% of all runners.  
```{r}
dfAgg = womens_table_T %>% filter(womens_table_T$AgeBin == "26-35") %>% group_by(year, AgeBin) %>% dplyr::summarise(AgeBinVolume = n())
dfAgg$TotalRunners = (womens_table_T %>% group_by(year) %>% dplyr::summarise(TotalRunners = n()))$TotalRunners
dfAgg$Proportion = dfAgg$AgeBinVolume / dfAgg$TotalRunners

ggplot(dfAgg, aes(x=year, y=Proportion)) +
  geom_col() +
  labs(title="Proportion of 26-35 year old runners by year") 
```

So to summarize, it's expected that 50% of the runners will be in the 26-35 year old age group in 2013. The median race time will be 97-98 minutes or approximately 9.7 minute pace.  



##########################Sabrina add on##########################################

When we consider how things have changed, we were asked to look at the average racer time.  For that, we are going to first consider a LOESS prediction.  Because this is a nonparametric method, and will weight itself relative to time, we feel like this should give indication of whether or not the average runner is getting faster or slower.
```{r}
#Predict LOESS on time
womens_table_T$RawTime_M = as.numeric(womens_table_T$RawTime_M) #need data to be numeric for LOESS
womens_table_T$year = as.numeric(womens_table_T$year)  #need data to be numeric for LOESS
plot(RawTime_M~year, ylim = c(94,100), data=womens_table_T, main="Year V Time")
out <- loess(RawTime_M~year, data=womens_table_T)
curve(predict(out, newdata=data.frame(year = x)), add=TRUE)
```
As shown above average time actually increased across the 14 year window, but did show a downturn in 2011 and 2012.  This indicates that the finishers may in fact be speeding back up again.

Our next objective is to determine if anywhere along the way the average time has sectioned off in a way that indicates multiple trends.  For that, we will run a changepoint evaluation on the average.
```{r}
#Run a piecewise fit next
plotdata = womens_table_T %>% 
 group_by(year)   %>% 
  summarise(average = mean(RawTime_M), .groups = "keep")

plotdata$average = as.integer(plotdata$average)
plotdata$average = as.ts(plotdata$average)

library(changepoint)
dis.bs <- cpt.meanvar(plotdata$average, test.stat = "Poisson", method = "BinSeg")
cpts.ts(dis.bs)
plot(dis.bs, cpt.width = 3)
```
You can see that this produced a single average.  This could be appropriately stating that nothing has statistically changed.  More likely, we do not have enough data points to find clean break points.