---
title: "Case_Study_2_Code"
author: "Brian Gaither, Sean Mcwhirter, Andrew Mejia, Sabrina Purvis"
date: "`r Sys.time()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 6
  github_document:
    toc: yes
    toc_depth: 6
  word_document:
    toc: yes
    toc_depth: '6'
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The following code has been adopted from the text book 
# NOLAN, D. (2017). DATA SCIENCE IN R: A case studies approach to computational reasoning and problem solving. Place of publication not identified: CRC Press.
# https://github.com/ngupta23/ds7333_qtw/tree/master/case_study_1/submission_Kannan_Moro_Gupta/code
# and class room starter code 
```{r}
library(tidyverse)
library(magrittr)
library(plotly)
library(DT)
library(flexclust)
```

## Clustering Distance and Mac Addresses of interest
## Since we are running this experiment over multiple loops, variable macAddressInt will be changed to include :c0 and :cd
```{r}
#Clustering Distance methods 
Cluster_Dist = 'Manhattan'
errMethod = ifelse(Cluster_Dist == 'Manhattan','MAE','RMSE')

#inverse of distance PG 42 
weightFormula = function(x){ 1/(x)} 

# Required MAC Addresses of interest 
macAddressInt = c(
  #'00:0f:a3:39:e1:c0', #Exclude from list
  '00:0f:a3:39:dd:cd',
  '00:14:bf:b1:97:8a',
  '00:14:bf:3b:c7:c6',
  '00:14:bf:b1:97:90',
  '00:14:bf:b1:97:8d',
  '00:14:bf:b1:97:81'
)

```

## This is the directory of the data f_paths passed out during class. 
```{r}
offline_f_path = "/home/andrew/Desktop/DS7333/offline.final.trace.txt"
#offline_f_path = "C:/Users/BGaither/OneDrive - McAfee/Other/school/SMU/Courses/Spring 2021/Quantifying The World/Case Study 2/Data/offline.final.trace.txt"
online_f_path = "/home/andrew/Desktop/DS7333/online.final.trace.txt"
#online_f_path = "C:/Users/BGaither/OneDrive - McAfee/Other/school/SMU/Courses/Spring 2021/Quantifying The World/Case Study 2/Data/online.final.trace.txt"
```

# The following function are from the starter code to process the raw data 

```{r}
# Create a function to parse the data based on the tokens passed 
processLine = function(x){
  # Split the line on: ';', '=' and ','
  tokens = strsplit(x, "[;=,]")[[1]]
  
  # The hand held device
  # is in the 1st 10 tokens (refer to text on page 9)
  # We change our function to return NULL if the tokens vector only has 10 elements (refer to page 11) as in the case of a missing signal
  if (length(tokens) == 10) {
    return(NULL)
  }
  
  # The tokens after the 10th one represent the signal strength at the access points (page 9). 
  # Split the tokens into individual attributes:  MAC address, Signal, Channel and Device Type
  # Device mode type 3 is important  page 6 and 7 table 1.1 
  # Device mode type 1 is the adhoc device
  # save this into tmp as a matrix with tokens 1:10 
  tmp = matrix(data = tokens[ - (1:10) ], ncol = 4, byrow = TRUE)
  
  #We can think of these as rows in a 4-column matrix or data frame giving the MAC address,signal, channel, and device type, 
  #So let’s unravel these and build a matrix from the values.
  #Then we can bind these columns with the values from the first 10 entries.
  cbind(matrix(tokens[c(2, 4, 6:8, 10)], nrow(tmp), 6, byrow = TRUE), tmp)
}
```

#From page 15, round orientation takes a parameter angles and rounds based on 45 degree increments and reorientates the angle and splits the near zero angles to map angles near 0 to 0 and near 360 to 360. 
```{r}
roundOrientation = function(angles) {
  refs = seq(0, by = 45, length = 9)
  q = sapply(angles, function(o) which.min(abs(o - refs)))
    c(refs[1:8], 0)[q]
  }
```

# Page 19 exercise to create readData() function

```{r}
# The purpose of this function is  to read the data f_path as described in page 19,  transform the parsed f_path to the dataframe for analysis
# f_path is the name of offline and online data to be read in
# macAddressInt The list of Mac addresses of interest, initialize at Null as a placeholder
# Return A dataframe df


readData = function(f_path, macAddressInt=NULL){
  # Read in the raw "offline" text from f_path
  txt = readLines(f_path)
  
#Process and input raw data
  
  # Parse the f_path lines and not include '#' 
  lines = txt[substr(txt, 1, 1) != "#" ]
  
  #apply processesline function to read lines and save as a temporary variable 
  tmp = lapply(lines, processLine)
  
  # Convert the tmp variable to a data frame df through binding rows 
  df = as.data.frame(do.call("rbind", tmp), stringsAsFactors = FALSE)

# Begin process to construction offline dataframe 
  
  # Assign column names to the offline dataframe from time, scanMac, posX, posY, posZ, orientation angle, mac, signal, channel and type
  names(df) = c(
    "time", "scanMac", "posX", "posY", "posZ",
    "orientation", "mac", "signal",
    "channel", "type"
  )
  
  #Cast the string variables from the data to numeric variables 
  
  numericVars = c("time", "posX", "posY", "posZ", "orientation", "signal")
  df[numericVars] = lapply(df[numericVars], as.numeric)
  
  # We will only keep required device types and remove the adhoc device (mode 1) refer to page 7 table 1.1
  df = df[df$type != 1, ]

  # We are only interested in the mac addresses of interest
  # We will filter the data frame to only include macs of interest and return the attributes of the macs of interest
  df = df[df$mac %in% macAddressInt, ]

  
  # Drop the scanMac and posZ attributes 
  
  #https://stackoverflow.com/questions/4605206/drop-data-frame-columns-by-name 
  #Max Ghenis Answer
  
  df = within(df, rm("scanMac", "posZ")) 

  # Apply round orientation function to df orientation to round angles 
  df$angle = roundOrientation(df$orientation)
  
  # Create posXY column 
  df$posXY = paste(df$posX, df$posY, sep = "-")

  return(df)
}
```

# Offline data

```{r}
numMacs = length(macAddressInt)
numMacs
```


```{r}
offline = readData(f_path = offline_f_path, macAddressInt = macAddressInt)
dim(offline)
length(unique(offline$posXY))
```

# Online Data
```{r}
online = readData(f_path = online_f_path, macAddressInt = macAddressInt)
dim(online)
length(unique(online$posXY))
```

# Function (Reshape) page 33
```{r}
# Unlike text example, we will initialize sampleAngle as false, so we do not select a sample of angles, if True, the ans x will take a random sample from the 45 degree angles increments (pg 38). This sampling technique was borrowed from page 38. 

reshapeSS = function(data, varSignal = "signal", keepVars = c("posXY", "posX", "posY"), sampleAngle = FALSE) {
  refs = seq(0, by = 45, length = 8)
  byLocation =
  with(
    data,
    by(
      data,
      list(posXY),
      function(x) {
        #page 38 to select one angle at random for each location
        if (sampleAngle) x = x[x$angle == sample(refs, size = 1), ]
        ans = x[1, keepVars]
        #page 38, we impute from the the mean the missing values 
        avgSS = tapply(x[ , varSignal ], x$mac, mean)
        y = matrix(avgSS, nrow = 1, ncol = numMacs,
        dimnames = list(ans$posXY, names(avgSS)))
        cbind(ans, y)
      }
    )
  )
  newDataSS = do.call("rbind", byLocation)
  return(newDataSS)
}

```

# Reshape Test Data Keeping only the flowing variables, "posXY", "posX","posY", "orientation", "angle" 
```{r}
keepVars = c("posXY", "posX","posY", "orientation", "angle")
onlineSummary = reshapeSS(data = online, varSignal = "signal", keepVars = keepVars)
head(onlineSummary,10)
```


# Function Select Train Data page 33 - 34

```{r}
# Select Train Data selects the appropriate observations (based on test data orientation) from the original tall form data
# and reformats for training the KNN algorithm
# angleNewObs Angle  of the new observation
# Train_data Original format of offlineSummary (Tall form) if signals
# m the angles to include from signals 
# Return a dataframe for training KNN algorithm 

selectTrain = function(angleNewObs, train_data, m){
  
#Apply angle rounding 
  
  nearestAngle = roundOrientation(angles = angleNewObs)
  

# if angle has a remainder, of 1 round down to the nearest 45 degree angle else round up. 
  if (m %% 2 == 1) {
    angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
  } else {
    m = m + 1
    angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
    if (sign(angleNewObs - nearestAngle) > -1)
      angles = angles[ -1 ]
    else
      angles = angles[ -m ]
  }
  
  angles = angles + nearestAngle
  angles[angles < 0] = angles[ angles < 0 ] + 360
  angles[angles > 360] = angles[ angles > 360 ] - 360
  
  # Subset only those angles from original data (tall-skinny)
  train_data_subset = train_data[train_data$angle %in% angles, ]
  
  # Convert to Wide and average the data for the same positions 
  train_data_subset = reshapeSS(data = train_data_subset, varSignal = "signal")
  
  return(train_data_subset)
}
```

#Calculate Nearest Neighboors 
# Page 35 and 36 

```{r}
# Computes the distance of the new signal (single observation) to each observation in the training dataset based on Euclidian or Manhattan distance
# newSignal is the signal value from the validation data set observation
# trainSubset the training data to be passed 
# Weighted default is FALSE to include weighted formula if TRUE else if False, weighted will just repeat weight of 1 for the distances
# Returns dataframe containing same number of rows as that in the training data.
# To calculate distance 
#https://machinelearningmastery.com/distance-measures-for-machine-learning/#:~:text=Manhattan%20distance%20is%20calculated%20as,differences%20between%20the%20two%20vectors.&text=The%20Manhattan%20distance%20is%20related,and%20mean%20absolute%20error%20metric.


findNN = function(newSignal, trainSubset, weighted=FALSE, method = Cluster_Dist) {
  #Instead of fixing the range of the columns of the vector from 4 to 9, we will use numMacs - 1 as the column indexer
  diffs = apply(trainSubset[ , 4:(4+numMacs-1)], 1, function(x) x - newSignal)
  # The following code is credited to https://github.com/ngupta23/ds7333_qtw/tree/master/case_study_1/submission_Kannan_Moro_Gupta/code
  #If Euclidian distance, distance is the Mean Square Error 
  if(method=='Euclidian')  dists = apply(diffs, 2, function(x) sqrt(sum(x^2)) ) 
  #If Manhattan distance is the sum of the absolute error of the measurement 
  if(method=='Manhattan')  dists = apply(diffs, 2, function(x) sum(abs(x)) ) 
  #Sort the distances in ascending order 
  closest = order(dists)
  #only return closest distances 
  ordered_dist = dists[closest]
  #When true, return the ordered ascending closest neighbors with weight
  if(weighted == TRUE){
    weight = weightFormula(ordered_dist)
  }
  #If false, return all equal weights of 1
  if(weighted == FALSE){
    weight = rep(1, length(dists))
  }
  return(cbind(trainSubset[closest, 1:3], ordered_dist, weight))
}
```

#Page 35
# The following code is adopted and credited to https://github.com/ngupta23/ds7333_qtw/tree/master/case_study_1/submission_Kannan_Moro_Gupta/code

```{r}
# XY Prediction for a single value of k neighboors
# newSignals the signal values for the validation data for each observation
# newAngles the orientation of the validation data for each observation
# trainData the training data to be used
# numAngles number of closest reference angles to include in the data
# k  predicton for num neighbors = k
# weighted default FALSE, include weighted distance
# Returns dataframe with n rows  = number of (validation) observations and n columns = 2 
# Each row is the  prediction of the mean X and Y values for that observation
predXY = function(newSignals, newAngles, trainData, numAngles = 1, k = 3, weighted=FALSE){
  closeXY = list(length = nrow(newSignals))
  for (i in 1:nrow(newSignals)) {
    trainSS = selectTrain(newAngles[i], trainData, m = numAngles)
    closeXY[[i]] = findNN(
      newSignal = as.numeric(newSignals[i, ]),
      trainSubset = trainSS,
      weighted = weighted
    )
  }
  #PG 40, not use CUMSUM function, use weighted mean
  # Returns the k-means observation based on weights for a single observation 
  # df Dataframe containing 5 columns , XY, X, Y, Distance, Inverse Distance
  # k sorted nearest neighboors list to calculate weight 
  # A pair of XY weighted means values for k number of neighbors
  
  #Calculate the weigheted mean 
  #https://www.mathsisfun.com/data/weighted-mean.html
  # The following code is adopted and credited to https://github.com/ngupta23/ds7333_qtw/tree/master/case_study_1/submission_Kannan_Moro_Gupta/code
  k_means_single_obs = function(df, k){
    weights = df[1:k, 5]
    weighted_x = weighted.mean(df[1:k,2],weights)
    weighted_y = weighted.mean(df[1:k,3],weights)
    
    return(c(weighted_x, weighted_y))
  }
  
  estXY = lapply(closeXY, k_means_single_obs, k)
  estXY = do.call("rbind", estXY)
  return(estXY)
}
```

```{r}
calcError = function(estXY, actualXY, method = Cluster_Dist){
  if('numeric' %in% class(estXY)) rows = 1 else rows = nrow(estXY)
  if(method == 'Euclidean')  error = sqrt(sum(rowSums((estXY - actualXY)^2)))/rows
  if(method == 'Manhattan')  error = sum(rowSums(abs(estXY - actualXY)))/rows
  return(error)
}
```


# K-Fold 

## Setup

#Page 37
```{r}
#subset size
set.seed(959)
v = 11
permuteLocs = sample(unique(offline$posXY))
permuteLocs = matrix(permuteLocs, ncol = v, nrow = floor(length(permuteLocs)/v))
permuteLocs
```


The following code will investigate our three MAC scenarios. 1) Drop 00:0f:a3:39:e1:c0 and keep the rest, 2) Drop 00:0f:a3:39:dd:cd and keep the rest, and 3) Keep all MAC addresses.  We'll also evaluate whether looking for up to 3 angles for prediction purposes aids in finding a better prediction.
```{r}
#Iterate through the MAC scenarios and generate the elbow plot while also checking for the best number of angles

K=20
noAngles = 3
v == 11
macPerms = c("Drop: 00:0f:a3:39:e1:c0","Drop: 00:0f:a3:39:dd:cd","Keep All")
# Required MAC Addresses of interest, here we include all mac addresses 
macAddressInt = c(
  '00:0f:a3:39:e1:c0', #Exclude from list
  '00:0f:a3:39:dd:cd',
  '00:14:bf:b1:97:8a',
  '00:14:bf:3b:c7:c6',
  '00:14:bf:b1:97:90',
  '00:14:bf:b1:97:8d',
  '00:14:bf:b1:97:81'
)
numMacs = length(macAddressInt)
#read the data with all mac addresses, we'll subset to remove specific mac addresses later
offline = readData(f_path = offline_f_path, macAddressInt = macAddressInt)
online = readData(f_path = online_f_path, macAddressInt = macAddressInt)

#set mac addresses of interest
macAddressIteration = macAddressInt
numMacs = length(macAddressIteration)
#used in the onlineFold creation
keepVars = c("posXY", "posX","posY", "orientation", "angle")

#create dataframe to hold results with errors
dfResults <- data.frame("Scenario"=factor(),"AngleCount"=integer(),"K_value"=integer(),"Error"=double())

for (f in 1:length(macPerms)){
  #in this loop we need to make sure our data only consists of the mac addresses we care about
  if (f < 3){
    #the first two scenarios involve dropping a mac address, this logic handles that
    print(paste0("Current Mac Scenario: ", f))
    macAddressIteration = macAddressInt[-f]
    numMacs = length(macAddressIteration)
  }else{
    print(paste0("Current Mac Scenario: ", f))
    macAddressIteration = macAddressInt
    numMacs = length(macAddressIteration)
  }
  offlineIt = subset(offline, mac %in% macAddressIteration)
  onlineIt = subset(online, mac %in% macAddressIteration)
  onlineCVSummary = reshapeSS(offlineIt, keepVars = keepVars, sampleAngle = TRUE)
  for (a in 1:noAngles){
    print(paste0("     Current number of angles: ", a))
    for (j in 1:v){
      #create folds for cross validation  
      print(paste0("          Current crossfold: ", j))
      onlineFold = subset(onlineCVSummary, posXY %in% permuteLocs[,j])
      if(f==3){print(paste0("          onlinefold created: ", j))}
      offlineFold = subset(offlineIt, posXY %in% permuteLocs[,-j])
      if(f==3){print(paste0("          offlinefold created: ", j))}
      actualFold = onlineFold[, c("posX","posY")]
      if(f==3){print(paste0("          actualfold created: ", j))}

      for (i in 1:K){
          if(f<3){
            estFold = predXY(newSignals = onlineFold[, 6:11], newAngles = onlineFold[, 4], offlineFold, numAngles = a, k = i)
          }else{
            estFold = predXY(newSignals = onlineFold[, 6:12], newAngles = onlineFold[, 4], offlineFold, numAngles = a, k = i)
          }
          myError = calcError(estFold, actualFold)
          print(paste0("               estimation and error calculated: ", i))
          if((f==1)&(a==1)&(j==1)&(i==1)){
            print(paste0("                    first time loading results K: ", i))
            dfResults <- data.frame("Scenario"=macPerms[f], "AngleCount"=a, "K_value"=i,"Error"=myError)
          }else{
            print(paste0("                    subsequent times loading results K: ", i))
            dfRow <- data.frame("Scenario"=macPerms[f], "AngleCount"=a, "K_value"=i,"Error"=myError)
            dfResults <- rbind(dfResults, dfRow)
          }
        
        }

      }
      
    }
}


```


```{r}
#write.csv(dfResults,"C:/Users/BGaither/OneDrive - McAfee/Other/school/SMU/Courses/Spring 2021/Quantifying The World/Case Study 2/Data/cv_results.csv", row.names = FALSE)
write.csv(dfResults,"/home/andrew/Desktop/DS7333/cv_results.csv", row.names = FALSE)
dfResults
```

```{r}

```

```{r}
library(ggplot2)
#dfResults = read.csv("C:/Users/BGaither/OneDrive - McAfee/Other/school/SMU/Courses/Spring 2021/Quantifying The World/Case Study 2/Data/cv_results.csv",header = TRUE)
#dfResGroup = dfResults %>% filter(AngleCount==3) %>% group_by(Scenario, AngleCount, K_value) %>% summarize(Error_Avg = mean(Error))
dfResults$AngleCount <- as.factor(dfResults$AngleCount)
dfResGroup = dfResults %>% group_by(Scenario, AngleCount, K_value) %>% summarize(Error_Avg = mean(Error))
#Individual aes() …. an aesthetic for each layer
ggplot(data = dfResGroup) + 
geom_line(mapping = aes(x=K_value, y=Error_Avg, color=AngleCount), size=1.2) + facet_grid(rows=vars(AngleCount), cols=vars(Scenario)) + theme_classic() + labs(title="Error Elbow Curve by Scenario")
```

```{r}
dfResultsGroupMin = dfResGroup %>% group_by(Scenario, AngleCount) %>% summarize(Error_Avg = min(Error_Avg))
dfOptK = merge(dfResGroup, dfResultsGroupMin, by=c("Scenario", "AngleCount", "Error_Avg"))
dfOptK
```

```{r}
ggplot(dfResults) + geom_boxplot(mapping = aes(x=Scenario, y=Error)) + facet_grid(rows=vars(AngleCount)) + theme_classic() + labs(title="Boxplot of Error by Scenario by Numer of Angles")
```
From the boxplot, we can see that there are some extreme error values when dropping 00:0f:a3:39:dd:cd.  Let's check dropping 00:0f:a3:39:e1:c0 and do an end to end analysis.
```{r}

macAddressInt = c(
  #'00:0f:a3:39:e1:c0', #Exclude from list
  '00:0f:a3:39:dd:cd',
  '00:14:bf:b1:97:8a',
  '00:14:bf:3b:c7:c6',
  '00:14:bf:b1:97:90',
  '00:14:bf:b1:97:8d',
  '00:14:bf:b1:97:81'
)
numMacs = length(macAddressInt)

offlineIt = subset(offline, mac %in% macAddressInt)
onlineIt = subset(online, mac %in% macAddressInt)
keepVars = c("posXY", "posX","posY", "orientation", "angle")
onlineCVSummary = reshapeSS(offlineIt, keepVars = keepVars, sampleAngle = TRUE)

i=6
noAngles = 1

actualLocation = onlineCVSummary[ , c("posX", "posY")]
estLocation = predXY(newSignals = onlineCVSummary[ , 6:11], newAngles = onlineCVSummary[ , 4], offlineIt, numAngles = noAngles, k = i)
locError =  calcError(estLocation, actualLocation)

locError
```
Let's check the actual locations and then we'll compare the predicted locations for spot check.
```{r}
actualLocation
```

Below we have the predicted locations, we do see that most predictions are in the ball park, however it's not perfect.
```{r}
dfEstLoc = as.data.frame(estLocation)
names(dfEstLoc)[names(dfEstLoc)=="V1"] <-"posX"
names(dfEstLoc)[names(dfEstLoc)=="V2"] <-"posY"
dfEstLoc
```

Let's take a look at what the predicted versus actual locations look like plotted out.
```{r}
ggplot(actualLocation,aes(posX,posY)) + 
geom_point(size=1.2) + geom_point(data=dfEstLoc, color="blue")+ theme_classic() + labs(title="Actual Location versus Predicted Location")
```

Now, we will reduce it to just 5 observations to get a closer look.
```{r}
ggplot(head(actualLocation),aes(posX,posY)) + 
geom_point(size=1.2) + geom_point(data=head(dfEstLoc), color="blue")+ theme_classic() + labs(title="Actual Location versus Predicted Location for 5 Observations")+ xlim(0, 35)
```

Now, the next best scenario appears to be including 3 angles and keeping all MAC addresses. Let's take this through an end to end analysis.
From the boxplot, we can see that there are some extreme error values when dropping 00:0f:a3:39:dd:cd.  Let's check dropping 00:0f:a3:39:e1:c0 and do an end to end analysis.
```{r}

macAddressInt = c(
  '00:0f:a3:39:e1:c0', #Exclude from list
  '00:0f:a3:39:dd:cd',
  '00:14:bf:b1:97:8a',
  '00:14:bf:3b:c7:c6',
  '00:14:bf:b1:97:90',
  '00:14:bf:b1:97:8d',
  '00:14:bf:b1:97:81'
)
numMacs = length(macAddressInt)

offlineIt = subset(offline, mac %in% macAddressInt)
onlineIt = subset(online, mac %in% macAddressInt)
keepVars = c("posXY", "posX","posY", "orientation", "angle")
onlineCVSummary = reshapeSS(offlineIt, keepVars = keepVars, sampleAngle = TRUE)

i=6
noAngles = 1

actualLocation = onlineCVSummary[ , c("posX", "posY")]
estLocation = predXY(newSignals = onlineCVSummary[ , 6:12], newAngles = onlineCVSummary[ , 4], offlineIt, numAngles = noAngles, k = i)
locError =  calcError(estLocation, actualLocation)

locError
```

Let's check the actual locations and then we'll compare the predicted locations for spot check.
```{r}
actualLocation
```

Below we have the predicted locations, we do see that most predictions are in the ball park, however it's not perfect.
```{r}
dfEstLoc = as.data.frame(estLocation)
names(dfEstLoc)[names(dfEstLoc)=="V1"] <-"posX"
names(dfEstLoc)[names(dfEstLoc)=="V2"] <-"posY"
dfEstLoc
```

Let's take a look at what the predicted versus actual locations look like plotted out.
```{r}
ggplot(actualLocation,aes(posX,posY)) + 
geom_point(size=1.2) + geom_point(data=dfEstLoc, color="blue")+ theme_classic() + labs(title="Actual Location versus Predicted Location")
```

Now, we will reduce it to just 5 observations to get a closer look.
```{r}
ggplot(head(actualLocation),aes(posX,posY)) + 
geom_point(size=1.2) + geom_point(data=head(dfEstLoc), color="blue")+ theme_classic() + labs(title="Actual Location versus Predicted Location for 5 Observations")+ xlim(0, 35)
```

#Modeling Scenario (Extended Case)
Weighted KNN

#Calculate Nearest Neighboors (Weighted)
# Page 35 and 36 

```{r}
# Computes the distance of the new signal (single observation) to each observation in the training dataset based on Euclidian or Manhattan distance
# newSignal is the signal value from the validation data set observation
# trainSubset the training data to be passed 
# Weighted default is FALSE to include weighted formula if TRUE else if False, weighted will just repeat weight of 1 for the distances
# Returns dataframe containing same number of rows as that in the training data.
# To calculate distance 
#https://machinelearningmastery.com/distance-measures-for-machine-learning/#:~:text=Manhattan%20distance%20is%20calculated%20as,differences%20between%20the%20two%20vectors.&text=The%20Manhattan%20distance%20is%20related,and%20mean%20absolute%20error%20metric.


findNN = function(newSignal, trainSubset, weighted=TRUE, method = Cluster_Dist) {
  #Instead of fixing the range of the columns of the vector from 4 to 9, we will use numMacs - 1 as the column indexer
  diffs = apply(trainSubset[ , 4:(4+numMacs-1)], 1, function(x) x - newSignal)
  # The following code is credited to https://github.com/ngupta23/ds7333_qtw/tree/master/case_study_1/submission_Kannan_Moro_Gupta/code
  #If Euclidian distance, distance is the Mean Square Error 
  if(method=='Euclidian')  dists = apply(diffs, 2, function(x) sqrt(sum(x^2)) ) 
  #If Manhattan distance is the sum of the absolute error of the measurement 
  if(method=='Manhattan')  dists = apply(diffs, 2, function(x) sum(abs(x)) ) 
  #Sort the distances in ascending order 
  closest = order(dists)
  #only return closest distances 
  ordered_dist = dists[closest]
  #When true, return the ordered ascending closest neighbors with weight
  if(weighted == TRUE){
    weight = weightFormula(ordered_dist)
  }
  #If false, return all equal weights of 1
  if(weighted == FALSE){
    weight = rep(1, length(dists))
  }
  return(cbind(trainSubset[closest, 1:3], ordered_dist, weight))
}
```

#Page 35
Weighted Neighbors
# The following code is adopted and credited to https://github.com/ngupta23/ds7333_qtw/tree/master/case_study_1/submission_Kannan_Moro_Gupta/code

```{r}
# XY Prediction for a single value of k neighboors
# newSignals the signal values for the validation data for each observation
# newAngles the orientation of the validation data for each observation
# trainData the training data to be used
# numAngles number of closest reference angles to include in the data
# k  predicton for num neighbors = k
# weighted default FALSE, include weighted distance
# Returns dataframe with n rows  = number of (validation) observations and n columns = 2 
# Each row is the  prediction of the mean X and Y values for that observation
predXY = function(newSignals, newAngles, trainData, numAngles = 1, k = 3, weighted=TRUE){
  closeXY = list(length = nrow(newSignals))
  for (i in 1:nrow(newSignals)) {
    trainSS = selectTrain(newAngles[i], trainData, m = numAngles)
    closeXY[[i]] = findNN(
      newSignal = as.numeric(newSignals[i, ]),
      trainSubset = trainSS,
      weighted = weighted
    )
  }
  #PG 40, not use CUMSUM function, use weighted mean
  # Returns the k-means observation based on weights for a single observation 
  # df Dataframe containing 5 columns , XY, X, Y, Distance, Inverse Distance
  # k sorted nearest neighboors list to calculate weight 
  # A pair of XY weighted means values for k number of neighbors
  
  #Calculate the weigheted mean 
  #https://www.mathsisfun.com/data/weighted-mean.html
  # The following code is adopted and credited to https://github.com/ngupta23/ds7333_qtw/tree/master/case_study_1/submission_Kannan_Moro_Gupta/code
  k_means_single_obs = function(df, k){
    weights = df[1:k, 5]
    weighted_x = weighted.mean(df[1:k,2],weights)
    weighted_y = weighted.mean(df[1:k,3],weights)
    
    return(c(weighted_x, weighted_y))
  }
  
  estXY = lapply(closeXY, k_means_single_obs, k)
  estXY = do.call("rbind", estXY)
  return(estXY)
}
```

The following code will investigate our three MAC scenarios and weight neighbors based on distance. 1) Drop 00:0f:a3:39:e1:c0 and keep the rest, 2) Drop 00:0f:a3:39:dd:cd and keep the rest, and 3) Keep all MAC addresses.  We'll also evaluate whether looking for up to 3 angles for prediction purposes aids in finding a better prediction .
```{r}
#Iterate through the MAC scenarios and generate the elbow plot while also checking for the best number of angles
set.seed(959)
K=20
noAngles = 3
v == 11
macPerms = c("Drop: 00:0f:a3:39:e1:c0","Drop: 00:0f:a3:39:dd:cd","Keep All")
# Required MAC Addresses of interest, here we include all mac addresses 
macAddressInt = c(
  '00:0f:a3:39:e1:c0', #Exclude from list
  '00:0f:a3:39:dd:cd',
  '00:14:bf:b1:97:8a',
  '00:14:bf:3b:c7:c6',
  '00:14:bf:b1:97:90',
  '00:14:bf:b1:97:8d',
  '00:14:bf:b1:97:81'
)
numMacs = length(macAddressInt)
#read the data with all mac addresses, we'll subset to remove specific mac addresses later
offline = readData(f_path = offline_f_path, macAddressInt = macAddressInt)
online = readData(f_path = online_f_path, macAddressInt = macAddressInt)

#set mac addresses of interest
macAddressIteration = macAddressInt
numMacs = length(macAddressIteration)
#used in the onlineFold creation
keepVars = c("posXY", "posX","posY", "orientation", "angle")

#create dataframe to hold results with errors
dfResults <- data.frame("Scenario"=factor(),"AngleCount"=integer(),"K_value"=integer(),"Error"=double())

for (f in 1:length(macPerms)){
  #in this loop we need to make sure our data only consists of the mac addresses we care about
  if (f < 3){
    #the first two scenarios involve dropping a mac address, this logic handles that
    print(paste0("Current Mac Scenario: ", f))
    macAddressIteration = macAddressInt[-f]
    numMacs = length(macAddressIteration)
  }else{
    print(paste0("Current Mac Scenario: ", f))
    macAddressIteration = macAddressInt
    numMacs = length(macAddressIteration)
  }
  offlineIt = subset(offline, mac %in% macAddressIteration)
  onlineIt = subset(online, mac %in% macAddressIteration)
  onlineCVSummary = reshapeSS(offlineIt, keepVars = keepVars, sampleAngle = TRUE)
  for (a in 1:noAngles){
    print(paste0("     Current number of angles: ", a))
    for (j in 1:v){
      #create folds for cross validation  
      print(paste0("          Current crossfold: ", j))
      onlineFold = subset(onlineCVSummary, posXY %in% permuteLocs[,j])
      if(f==3){print(paste0("          onlinefold created: ", j))}
      offlineFold = subset(offlineIt, posXY %in% permuteLocs[,-j])
      if(f==3){print(paste0("          offlinefold created: ", j))}
      actualFold = onlineFold[, c("posX","posY")]
      if(f==3){print(paste0("          actualfold created: ", j))}

      for (i in 1:K){
          if(f<3){
            estFold = predXY(newSignals = onlineFold[, 6:11], newAngles = onlineFold[, 4], offlineFold, numAngles = a, k = i)
          }else{
            estFold = predXY(newSignals = onlineFold[, 6:12], newAngles = onlineFold[, 4], offlineFold, numAngles = a, k = i)
          }
          myError = calcError(estFold, actualFold)
          print(paste0("               estimation and error calculated: ", i))
          if((f==1)&(a==1)&(j==1)&(i==1)){
            print(paste0("                    first time loading results K: ", i))
            dfResults <- data.frame("Scenario"=macPerms[f], "AngleCount"=a, "K_value"=i,"Error"=myError)
          }else{
            print(paste0("                    subsequent times loading results K: ", i))
            dfRow <- data.frame("Scenario"=macPerms[f], "AngleCount"=a, "K_value"=i,"Error"=myError)
            dfResults <- rbind(dfResults, dfRow)
          }
        
        }

      }
      
    }
}


```

```{r}
#write.csv(dfResults,"C:/Users/BGaither/OneDrive - McAfee/Other/school/SMU/Courses/Spring 2021/Quantifying The World/Case Study 2/Data/cv_results.csv", row.names = FALSE)
dfResults
```

#Elbow plots to identify best k-value
```{r}
library(ggplot2)
#dfResults = read.csv("C:/Users/BGaither/OneDrive - McAfee/Other/school/SMU/Courses/Spring 2021/Quantifying The World/Case Study 2/Data/cv_results.csv",header = TRUE)
#dfResGroup = dfResults %>% filter(AngleCount==3) %>% group_by(Scenario, AngleCount, K_value) %>% summarize(Error_Avg = mean(Error))
dfResults$AngleCount <- as.factor(dfResults$AngleCount)
dfResGroup = dfResults %>% group_by(Scenario, AngleCount, K_value) %>% summarize(Error_Avg = mean(Error))
#Individual aes() …. an aesthetic for each layer
ggplot(data = dfResGroup) + 
geom_line(mapping = aes(x=K_value, y=Error_Avg, color=AngleCount), size=1.2) + facet_grid(rows=vars(AngleCount), cols=vars(Scenario)) + theme_classic() + labs(title="Error Elbow Curve by Scenario")
```

#Top combinations
```{r}
dfResultsGroupMin = dfResGroup %>% group_by(Scenario, AngleCount) %>% summarize(Error_Avg = min(Error_Avg))
dfOptK = merge(dfResGroup, dfResultsGroupMin, by=c("Scenario", "AngleCount", "Error_Avg"))
dfOptK
```

#Box Plots
```{r}
ggplot(dfResults) + geom_boxplot(mapping = aes(x=Scenario, y=Error)) + facet_grid(rows=vars(AngleCount)) + theme_classic() + labs(title="Boxplot of Error by Scenario by Number of Angles")
```

#Analysis
```{r}

macAddressInt = c(
  #'00:0f:a3:39:e1:c0', #Exclude from list
  '00:0f:a3:39:dd:cd',
  '00:14:bf:b1:97:8a',
  '00:14:bf:3b:c7:c6',
  '00:14:bf:b1:97:90',
  '00:14:bf:b1:97:8d',
  '00:14:bf:b1:97:81'
)
numMacs = length(macAddressInt)

offlineIt = subset(offline, mac %in% macAddressInt)
onlineIt = subset(online, mac %in% macAddressInt)
keepVars = c("posXY", "posX","posY", "orientation", "angle")
onlineCVSummary = reshapeSS(offlineIt, keepVars = keepVars, sampleAngle = TRUE)

i=4
noAngles = 3

actualLocation = onlineCVSummary[ , c("posX", "posY")]
estLocation = predXY(newSignals = onlineCVSummary[ , 6:11], newAngles = onlineCVSummary[ , 4], offlineIt, numAngles = noAngles, k = i)
locError =  calcError(estLocation, actualLocation)

locError
```

Let's check the actual locations and then we'll compare the predicted locations for spot check.
```{r}
actualLocation
```

Below we have the predicted locations, we do see that most predictions are in the ball park, however it's not perfect.
```{r}
dfEstLoc = as.data.frame(estLocation)
names(dfEstLoc)[names(dfEstLoc)=="V1"] <-"posX"
names(dfEstLoc)[names(dfEstLoc)=="V2"] <-"posY"
dfEstLoc
```

Let's take a look at what the predicted versus actual locations look like plotted out.
```{r}
ggplot(actualLocation,aes(posX,posY)) + 
geom_point(size=1.2) + geom_point(data=dfEstLoc, color="blue")+ theme_classic() + labs(title="Actual Location versus Predicted Location")
```

Now, we will reduce it to just 5 observations to get a closer look.
```{r}
ggplot(head(actualLocation),aes(posX,posY)) + 
geom_point(size=1.2) + geom_point(data=head(dfEstLoc), color="blue")+ theme_classic() + labs(title="Actual Location versus Predicted Location for 5 Observations")+ xlim(0, 35)
```




#Conclusion
The conclusion of this analysis is dropping the MAC address 00:0f:a3:39:e1:c0 and using only 1 angle with an optimal k value of 6. This is recommended as it’s the simplest approach and results in the most accurate prediction.   